\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
% %% complete the rights form.
% \setcopyright{acmlicensed}
% \copyrightyear{2026}
% \acmYear{2026}
% \acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[CS5220 '26]{Proposal}{March 27--03,
  2026}{Ithaca, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
% \acmISBN{978-1-4503-XXXX-X/2018/06}


%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}
\renewcommand{\abstractname}{Implementation Approach}

\settopmatter{printacmref=false} % Removes the ACM reference format block
\renewcommand\footnotetextcopyrightpermission[1]{} % Removes the permission footnote

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{CS 5220 Homework 4 - Distributed Matrix Multiplication on WSE-2}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Dalton Luce}
\authornote{NetID: dcl252; Perlmutter username: dcl252}
\email{dcl252@cornell.edu}
\affiliation{%
  \institution{Cornell University}
  \city{Ithaca}
  \state{NY}
  \country{United States}}


%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Luce}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}

My first step in implementing the matrix multiplication algorithm was to establish
a working column broadcast. I established the necessary layout for the top row to
broadcast the columns of \texttt{B} to the rest of the grid. The hardest part of this was
establishing colorswap on the \texttt{RAMP} for the top row (such that when the top row
broadcasts on \texttt{broadcast\_rx\_color}, it swaps to \texttt{broadcast\_tx\_color} on the \texttt{RAMP} and
is sent correctly down the column). This required setting swap on the y and x
directions, which was hard to find within the documentation. I established my top
row broadcast was working by checking the sim.log to see if all rows in the same
column were receiving the same data from the top row.

Once I had this working, I extended the column broadcast to the rest of the grid.
I was a bit confused by the hint for this part, which implied that a \texttt{SWITCH\_ADV}
control wavelet was needed for signaling the next row to start broadcasting. I 
found that this was not necessary, and that I could simply maintain the same
switch state for all PEs on \texttt{broadcast\_tx\_color}. Instead, I sent a \texttt{NOP} control signal
(meaning no changes to the switch needs to be applied). However, this control signal
can still trigger a task on the receiving PE's compute engine (CE), which is 
exactly what we want. So, to trigger the next row to start broadcasting, I simply
send a control wavelet with two \texttt{NOP} control signals. The first one is popped by 
the current (sending) row (there is a mask that indicates the first \texttt{NOP} does not trigger a
CE task). The control wavelet is then routed to the next row, where the second
\texttt{NOP} is popped. This \texttt{NOP} does trigger a CE task, which starts the next row's
broadcast. Once again, I established this was working by checking the sim.log to
see if all rows in the same column were receiving the same data from the top row,
and that the next row started broadcasting after the previous row finished.

Once I had the column broadcast working, I thought about how to transpose the matrices
\texttt{A}, \texttt{B}, and \texttt{C} to make DSD access easy and fast. Much like
from Homework 1, a fast DSD access pattern is one that is contiguous in memory
(no strides created from multiplication). With the advise of Google Gemini, and
working through the algorithm on paper, I finalized on a layout, found in
\texttt{config.py}.

Once the DSD access pattern was established, I implemented the SAXPY operation
for the inner loop of the matrix multiplication. This was pretty straightforward,
as it is a single builtin operation on the WSE-2 (\texttt{@fmacs}). I did forget
to advance A to the next column for the next incoming wavelet, which caused bugs
in the beginning, but was an easy fix once I realized the mistake.

Finally, I implemented the row reduction to sum up the partial results from each
row. I implemented the correct layout after understanding how the flyover color
should function (I misinterpreted the diagram from the recitation at first). I
initially computed the correct state on each reduction using modulo and division
operations. However, I realized that this was not necessary, and that I could
simply maintain the state in an integer on each PE, updating it with each
incoming wavelet. This proved to be much faster and cut down on tens of thousands 
of cycles from the reduction phase.

Overall, I ran into many bugs along the way. In addition to the staged implementation
approach detailed above, the best technique I found for debugging was to maximize
the visibility of the algorithm's state in the sim.log. This would not only help
my understanding of how the algorithm was progressing, but also greatly assist
LLMs in helping me debug.

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%


%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.


%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\vspace{-0.5em}
\section{Performance Analysis}


\vspace{-1.5em}

\begin{table}[h!]
\centering
\begin{tabular}{|l|r|}
\hline
\textbf{Parameter} & \textbf{Value} \\
\hline
sizeX & 8 \\
sizeY & 8 \\
M     & 128 \\
H     & 256 \\
N     & 64 \\
Total cycles & 218,606 \\
Idle cycles  & 16,877 \\
Runtime      & 201,729 \\
TSC Min      & 115,743 \\
TSC Max      & 115,895 \\
TSC Mean     & 115,821.9 \\
\hline
\end{tabular}
\caption{Performance statistics emitted by \texttt{test\_perf.sh}}
\label{tab:test_perf}
\end{table}

\vspace{-2.5em}

The three biggest optimizations I implemented were: 

\begin{itemize}
  \item Removing print statements from my implementation
  \item Tracking the reduction state machine with an integer on each PE instead
  of computing it with modulo and division operations
  \item zeroing out the \texttt{red\_in} buffer using a DSD \texttt{@fmovs} operation
 instead of a manual for loop
\end{itemize}

\vspace{-0.5em}
\section{Collaboration and GenAI Disclosure}

This homework was completed solely by me. I did not collaborate with any other
students in the class.

I used GenAI tools more on this assignment than previous assignments, as I
found them to be very helpful for understanding the WSE-2 architecture and
debugging my implementation. I used Google Gemini to help me understand the
documentation and come up with an initial layout for the matrices. I also used
Gemini to help me debug the many bugs I ran into along the way. The implementation
is my own, as the LLMs struggled with the syntax and semantics of the WSE-2, but
the LLMs were very helpful for debugging and understanding the algorithm. All
writing in this report is my own.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
\endinput
%%
%% End of file `sample-sigconf-authordraft.tex'.
